A Comparison of Parallel Graph Processing Benchmarks

### 2017-02-09 papers/publication/paper-pollard.tex
Renamed and refactored paper to be in LLNCS format
The abstract is almost ready for submission but the rest of the
paper is in a rough spot.

Updated the makefile and added the .cls files.

Still need to get power/energy usage and get more platforms and
algorithms and get more architectures and write the paper.

\documentclass[conference]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{hyperref}
\usepackage{pgfplotstable}
\usepackage{array} % Required for pgfplotstable dec sep align
\pgfplotsset{compat=1.12}
\usepackage{booktabs}

@@

\title{A Comparison of Parallel Graph Processing Benchmarks}
\author{
	\IEEEauthorblockN{Sam D. Pollard}
	\IEEEauthorblockA{Department of Computer and Information Science \\
		University of Oregon \\
		Eugene, OR, USA \\
		Email: spollard@cs.uoregon.edu
	}
	\and
	\IEEEauthorblockN{Boyana Norris}
	\IEEEauthorblockA{Department of Computer and Information Science \\
	University of Oregon \\
	Eugene, OR, USA \\
	Email: norris@cs.uoregon.edu
	}
}

\begin{document}
\maketitle
\begin{abstract}
There are an intimidating number of parallel graph processing systems. One survey in 2014 identified over 80. Since then, the landscape has changed; some packages have become inactive, even more projects have been introducted, and the previous top performers are being usurped. To address the high churn of projects and the complexity of selecting a system for a given problem, we analyze the performance and scalability of four parallel open source graph processing systems (GraphMat, GAP Benchmark Suite, GraphBIG, and PowerGraph) using synthetic and real-world datasets. We perform analysis for three algorithms: breadth first search, single source shortest paths, and PageRank. We compare our results to largest parallel graph processing performance and benchmarking project: Graphalytics. This paper examines previously overlooked aspects of parallel graph processing performance such as energy usage in addition to a detailed performance analysis modeled after the Graph500 that takes into consideration multiple runs and the phases of computation that graph programs typically undergo.
\end{abstract}

\section{Introduction}

@@

Graphalytics \cite{Capota:2015:Graphalytics} also attempts to remedy these problems. However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must provide Java methods to call the particular implementations and initialization shell scripts to satisfy dependencies and set environment variables. These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API.

@@

This report explores four shared memory parallel graph processing platforms. The first three are so-called ``reference implementations'' while the fourth is included because it has been cited as highly performant \cite{Sundaram:2015:GraphMat}. Our target is shared memory CPU processing. Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations for all algorithms. The systems are:

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/bfs_time.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis is logarithmic.}
	\label{fig:bfs-time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/bfs_dsc.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
	\label{fig:bfs-dsc}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/sssp_time.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis is logarithmic.}
	\label{fig:sssp-time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/pr_time.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis is logarithmic. GraphMat continues to run until none of the vertices' ranks change. [There is only one data point]}
	\label{fig:pr-time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/pr_iters.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis \emph{not} logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.}
	\label{fig:pr-iters}
\end{figure}
[Note that for figure Fig.\ref{fig:pr-iters}, mention time per iteration]

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/bfs_ss.pdf}
	\vspace{-18pt}
	\caption{Strong Scaling for $2^{20}$ vertices and an average of 16 edges per vertex.}
	\label{fig:bfs-ss}
\end{figure}

\begin{figure}
	\centering
	% trim left bottom right top
	\includegraphics[width=\columnwidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
	\caption{Speedup.}
	\label{fig:bfs-speedup}
\end{figure}

@@

\bibliographystyle{IEEEtranS}
\bibliography{../drp}
\end{document}

>>>>

\documentclass{llncs}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{hyperref}
\usepackage{cite} % Multiple citations in one \cite command
\usepackage{pgfplotstable}
\usepackage{array} % Required for pgfplotstable dec sep align
\pgfplotsset{compat=1.12}
\usepackage{booktabs}

@@

\begin{document}
\title{A Comparison of Parallel Graph Processing Benchmarks}
\author{Samuel D. Pollard \and Boyana Norris}
\institute{University of Oregon \\
	Eugene OR 97403, USA \\
	\email{\{spollard,norris\}@cs.uoregon.edu}
}
\maketitle
\begin{abstract}
There are an intimidating number of parallel graph processing systems. One survey in 2014 identified over 80. Since then, the landscape has changed; some packages have become inactive, even more have been introduced, and previous top performers have been usurped. To address the high churn of projects and the complexity of selecting a system for a given problem, we analyze the performance and scalability of four parallel open source graph processing systems (GraphMat, GAP Benchmark Suite, GraphBIG, and PowerGraph) using synthetic and real-world datasets. We perform analysis for three algorithms: breadth first search, single source shortest paths, and PageRank. We then compare our results to the most comprehensive parallel graph processing performance and benchmarking project: Graphalytics. We examine previously overlooked aspects of parallel graph processing performance such as energy usage in addition to a performance analysis model based on the Graph500.
\end{abstract}

\section{Introduction and Related Work}

@@

Graphalytics \cite{Capota:2015:Graphalytics} also attempts to remedy these problems. However, in order for Graphalytics to analyze a graph processing system---Graphalytics calls these platforms---a programmer must implement Java classes as wrappers to call the particular implementations, satisfy dependencies, and set environment variables. These all require some knowledge of the inner workings of each system in addition to familiarity with the Graphalytics API.

@@

This report explores four shared memory parallel graph processing platforms. The first three are so-called ``reference implementations'' while the fourth is included because it has been cited as highly performant \cite{Sundaram:2015:GraphMat}. Our target is shared memory CPU processing. Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} are not considered here because the authors do not provide reference implementations. The systems are:

@@

\begin{figure}
	\centering
	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_time.pdf}
	\caption{The $y$-axis is logarithmic.}
	\label{fig:bfs-time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_dsc.pdf}
	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
	\label{fig:bfs-dsc}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/sssp_time.pdf}
	\caption{The $y$-axis is logarithmic.}
	\label{fig:sssp-time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/pr_time.pdf}
	\caption{The $y$-axis is logarithmic. GraphMat continues to run until none of the vertices' ranks change. [There is only one data point]}
	\label{fig:pr-time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.6\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/pr_iters.pdf}
	\caption{The $y$-axis \emph{not} logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs until none of the vertices' ranks change.}
	\label{fig:pr-iters}
\end{figure}
[Note that for figure Fig.\ref{fig:pr-iters}, mention time per iteration]

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ss.pdf}
	\caption{Strong Scaling for $2^{20}$ vertices and an average of 16 edges per vertex.}
	\label{fig:bfs-ss}
\end{figure}

\begin{figure}
	\centering
	% trim left bottom right top
	\includegraphics[width=\columnwidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
	\caption{Speedup.}
	\label{fig:bfs-speedup}
\end{figure}

@@

\bibliographystyle{splncs03}
\bibliography{../drp}
\end{document}
